---
title: "HW3"
author: "Ophelia Li"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---
## Q1
### Read in the data of COVID-19 cases reported on 2020-10-18
```{r}
rm(list = ls())
library(tidyverse)
library(gtsummary)
library(faraway)
library(glmnet)
```

```{r}
county_count <- read_csv("10-18-2020.csv") %>%
  mutate(FIPS = as.numeric(FIPS)) %>%
  filter(Country_Region == "US") %>%
  print(width = Inf)
```

```{r}
names(county_count) <- str_to_lower(names(county_count))
```

```{r}
county_count %>%
  select(province_state) %>%
  distinct() %>%
  arrange(province_state) %>%
  print(n = Inf)
```

```{r}
county_count <- county_count %>%
  filter(!(province_state %in% c("Diamond Princess", "Grand Princess", 
                                 "Recovered", "Guam", "Northern Mariana Islands", 
                                 "Puerto Rico", "Virgin Islands"))) %>%
  print(width = Inf)
```

```{r}
county_count %>%
  # turn into long format for easy plotting
  pivot_longer(confirmed:recovered, 
               names_to = "case", 
               values_to = "count") %>%
  group_by(province_state) %>%
  ggplot() + 
  geom_col(mapping = aes(x = province_state, y = `count`, fill = `case`)) + 
  # scale_y_log10() + 
  labs(title = "US COVID-19 Situation on 2020-10-18", x = "State") + 
  theme(axis.text.x = element_text(angle = 90))
```

### Read in the 2020 county-level health ranking data
```{r}
county_info <- read_csv("health_rankings.csv") %>%
  filter(!is.na(county)) %>%
  # cast fips into dbl for use as a key for joining tables
  mutate(fips = as.numeric(fips)) %>%
  select(fips, 
         state,
         county,
         percent_fair_or_poor_health, 
         percent_smokers, 
         percent_adults_with_obesity, 
         # food_environment_index,
         percent_with_access_to_exercise_opportunities, 
         percent_excessive_drinking,
         # teen_birth_rate, 
         percent_uninsured,
         # primary_care_physicians_rate,
         # preventable_hospitalization_rate,
         # high_school_graduation_rate,
         percent_some_college,
         percent_unemployed,
         percent_children_in_poverty,
         # `80th_percentile_income`,
         # `20th_percentile_income`,
         percent_single_parent_households,
         # violent_crime_rate,
         percent_severe_housing_problems,
         overcrowding,
         # life_expectancy,
         # age_adjusted_death_rate,
         percent_adults_with_diabetes,
         # hiv_prevalence_rate,
         percent_food_insecure,
         # percent_limited_access_to_healthy_foods,
         percent_insufficient_sleep,
         percent_uninsured_2,
         median_household_income,
         average_traffic_volume_per_meter_of_major_roadways,
         percent_homeowners,
         # percent_severe_housing_cost_burden,
         population_2,
         percent_less_than_18_years_of_age,
         percent_65_and_over,
         percent_black,
         percent_asian,
         percent_hispanic,
         percent_female,
         percent_rural) %>%
  print(width = Inf)
```

```{r}
county_count <- county_count %>%
  filter(confirmed >= 5)
```

```{r}
county_data <- county_count %>%
  left_join(county_info, by = "fips") %>%
  print(width = Inf)
```

```{r}
summary(county_data)
```

```{r}
county_data %>%
  filter(is.na(state) & is.na(county)) %>%
  print(n = Inf)
```

```{r}
county_count %>%
  filter(is.na(fips)) %>%
  select(fips, admin2, province_state) %>%
  print(n = Inf)
```

```{r}
county_data <- county_count %>%
  # manually set FIPS for some counties
  mutate(fips = ifelse(admin2 == "Dukes and Nantucket" & province_state == "Massachusetts", 25019, fips)) %>% 
  mutate(fips = ifelse(admin2 == "Weber-Morgan" & province_state == "Utah", 49057, fips)) %>%
  filter(!(is.na(fips) | str_detect(admin2, "Out of") | str_detect(admin2, "Unassigned"))) %>%
  left_join(county_info, by = "fips") %>%
  drop_na() %>%
  print(width = Inf)
```

```{r}
summary(county_data)
```

```{r}
county_data <- county_data %>%
  mutate(state = as.factor(state)) %>%
  select(county, confirmed, deaths, state, percent_fair_or_poor_health:percent_rural)
summary(county_data)
```

```{r}
county_data %>%
  mutate(cfr = deaths / confirmed) %>%
  select(county, state, confirmed, deaths, cfr) %>%
  arrange(desc(cfr)) %>%
  top_n(10)
```

```{r}
write_csv(county_data, "covid19-county-data-20201018.csv.gz")
```

## Q2
Confirmed cases may die, thus uning deaths/confirmed is not a good predictor. And confirmed cases are not the number of diagnosed people.

## Q3
Logistic regression requires the observations to be independent of each other, but each individual may not be perfectly independent of each other in this study.

## Q4
```{r}
binaryreg=glm(cbind(deaths, confirmed - deaths) ~ . - county, family = binomial, data = county_data) 
summary(binaryreg)
```

```{r}
binaryreg %>%
  tbl_regression() %>%
  bold_labels() %>%
  bold_p(t = 0.01)
```

## Q5
Holding all other variables constant.

Interpret "percent_65_and_over" (old people may have a greater chance of dying from COVID-19): a one-unit increase in the percentage of people who are 65 years old and older increases the odds of the CFR by 4%.

Interpret "percent_adults_with_diabetes": a one-unit increase in the percentage of adults with diabetes increases the odds of the CFR by 1%.

Interpret "percent_children_in_poverty": a one-unit increase in the percentage of children in poverty increases the odds of the CFR by 2%.

## Q6
### analysis of deviance to the goodness of fit of the model
```{r}
pchisq(binaryreg$deviance, binaryreg$df.residual, lower = FALSE)
```
The result shows that the model does not fit well

### Analysis of deviance to the intercept-only model
```{r}
binaryreg$deviance
```

```{r}
binaryreg$df.residual
```

```{r}
binaryreg$null.deviance
```

```{r}
binaryreg$df.null
```

```{r}
1 - pchisq(binaryreg$null.deviance - binaryreg$deviance, binaryreg$df.null - binaryreg$df.residual)
```
The model is a better fit than the intercept-only model

## Q7
```{r}
moddev=drop1(binaryreg, test = "Chi")
rownames(moddev[order(moddev$`Pr(>Chi)`),][1:10,])
```

## Q8
```{r}
confint(binaryreg)
```

## Q9
```{r}
dev_resi=county_data %>%
  mutate(devres=residuals(binaryreg), linpred=predict(binaryreg)) %>%
  ggplot()+ 
  geom_point(mapping=aes(x=linpred, y=devres))+
  labs(x="linear predictor", y="deviance residuals")
print(dev_resi)
```

There are two potential outliers on the right of x=-6

```{r}
dev_resi+xlim(-6,-2)
```

When we take the two potential outliers out, the plot looks better now.

## Q10
```{r}
halfnorm(hatvalues(binaryreg), labs=county_data$county)
```

District of Columbia and Honolulu look like outliers. We will investigate more.

```{r}
county_data %>%
slice(c(309, 537)) %>%
  print(width=Inf)
```

It does not look like they are potential outliers in predictor space

## Q11
```{r}
modAIC=step(binaryreg, trace = FALSE)
summary(modAIC)
```

## Q12
```{r}
x_vars=model.matrix(cbind(deaths, confirmed - deaths) ~ . - 1 - county, 
                  data = county_data)
x_vars=x_vars[rep(1:nrow(county_data), each = 2), ]
y_vars=rep(c(1, 0), nrow(county_data))
combined=c(rbind(county_data$deaths, county_data$confirmed - county_data$deaths))
set.seed(6000)
cv_output=cv.glmnet(x_vars, y_vars, weights = combined, alpha = 1, 
                      family = "binomial", type.measure = "auc")
plot(cv_output)
```

```{r}
cv_output$lambda.min
```

```{r}
coef(cv_output, cv_output$lambda.min)
```